---
title: "Log-normal models"
author: "Stefano Coretta"
date: "27/04/2020"
output:
  html_document:
    highlight: tango
    number_sections: yes
    theme: paper
    toc: yes
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = here::here())
library(tidyverse)
library(ggrepel)
theme_set(theme_minimal())
library(brms)
library(bayesplot)
library(tidybayes)
library(broom.mixed)
# devtools::install_github("stefanocoretta/coretta2019eng@v1.0.0")
library(coretta2019eng) # Load Coretta 2019 data
data("eng_durations")
options(mc.cores = parallel::detectCores())
```

# Overview

Across a wide variety of languages, vowels tend to be shorter when followed by voiceless consonants, and longer when followed by voiced ones.
This phenomenon has been called the "voicing effect" [@mitleb1982] or "pre-fortis clipping" [@wells1990].
Among the earliest traceable mentions to this phenomenon there are @meyer1903 for English [cited in @lindblom1967], @meyer1904 for German, @meyer1909 for Hungarian, and @gregoire1911 for French [all cited in @maddieson1976].
After these, a great number of studies further confirmed the existence of the effect in these languages and reported it in an ever increasing list of others.

English is the language that by far received the most attention in relation to the voicing effect [@heffner1937; @house1953; @lisker1957; @zimmerman1958; @peterson1960; @house1961; @sharf1962; @sharf1964; @lindblom1967; @halle1967; @halle1967a; @slis1969a; @slis1969b; @chen1970; @klatt1973; @lisker1974; @raphael1975; @umeda1975; @javkin1976; @port1982; @mack1982; @luce1985; @van-summers1987; @kluender1988; @de-jong1991; @laeufer1992; @fowler1992; @de-jong2004; @warren2005; @ko2018; @glewwe2018; @sanker2019, among others].

In light of rising concerns for statistical power and publication bias [@roettger2018a; @kirby2018; @vasishth2019], it is fundamental to obtain a meta-analytical estimate of the voicing effect based on previous work, and to investigate potential biases that can affect the estimate.
A Bayesian meta-analysis was carried out on 17 Bayesian estimates of the effect of voicing on vowel duration obtained from a selection of 13 studies of English according to the method described in @nicenboim2018a.

# Read data

<!-- Specify which studies and how they were selected -->

```{r data, message=FALSE}
stressed <- read_csv("./data/raw/stressed.csv") %>%
  filter(manner == "stop") %>%
  mutate(
    voice = factor(voice, levels = c("voiceless", "voiced")),
    n_syl = factor(n_syl, levels = c("mono", "di", "tri")),
    word_pos_2 = ifelse(word_pos == "medial", "medial", "final")
  ) %>%
  mutate_if(is.character, as.factor)

studies <- stressed %>%
  select(study, n_syl:word_pos_2) %>%
  unique() %>%
  # Hussein 1994 is not included in the analysis, Port 1981 tri is included with
  # Port 1981 di. Klatt 1973 is not included since they only provided grand means.
  filter(study != "hussein1994", study != "port1981_tri", study != "klatt1973_mono", study != "klatt1973_di") %>%
  droplevels()

# ko18_data <- read_delim("./data/raw/ko2018.txt", delim = "\t")

iterations <- 4000

if (file.exists("./data/cache/estimated_ln_tbl.rds")) {
  estimated_ln_tbl <- readRDS("./data/cache/estimated_ln_tbl.rds")
}

if (file.exists("./data/cache/estimated_ln_sd_tbl.rds")) {
  estimated_ln_sd_tbl <- readRDS("./data/cache/estimated_ln_sd_tbl.rds")
}
```

# Get studies estimates

The first step is to get the estimate (with respective error) of the voicing effect for each study individually.
A log-normal distribution is used here as the distribution of the outcome variable (vowel duration), given evidence that the duration of speech intervals (segments, utterances, pauses) tend to be log-normally distributed [@rosen2005; @ratnikova2017; @gahl2019].
For each study, the calculated estimate is the ratio between vowel duration when the vowel is followed by a voiceless consonant vs when the following consonant is voiced (the "voicing effect (VE) ratio"; for example, a VE ratio of 1.5 means that vowels are 1.5 times longer when the following consonant is voiced).

Depending on the nature of the information available from each study, the following methods are used to obtain the estimated VE ratio:

- **Studies that provide measures of central tendency (means) and dispersion (standard deviations)**: Bayesian linear regressions with measurement error.
- **Studies that only provide measures of central tendency (means)**: Bayesian linear regression.
- **Studies that provide raw data**: Bayesian linear regression fitted to the raw data.

## Studies that provide means and standard deviation

Using measurement error models with studies that reported standard deviations together with the means.
The model includes `mi()` terms, where the standard error is calculated as $sd/\sqrt{N}$ ($N$ is the number of observations from which the standard deviation was obtained). See <https://github.com/paul-buerkner/brms/issues/643#issuecomment-483339742>.

```{r port1981}
p81_di <- filter(stressed, study == "port1981_di") %>%
  mutate(obs = 1:n())
p81_mono <- filter(stressed, study == "port1981_mono") %>%
  mutate(obs = 1:n())

# Fitting a lognormal likelihood with (voice|pair) leads to many divergent transitions.

priors <- c(
  prior(normal(0, 3.2), class = Intercept),
  prior(normal(0, 1), class = b, coef = voicevoiced),
  prior(normal(0, 0.1), class = sigma)
)

p81_di_ln_bm <- brm(
  v_duration | mi(sd/sqrt(n_obs)) ~
    voice,
  data = p81_di,
  prior = priors,
  family = lognormal,
  seed = 9899,
  file = "./data/cache/p81_di_ln_bm",
  control = list(adapt_delta = .9999, max_treedepth = 20),
  iter = iterations,
  save_mevars = TRUE
)

p81_mono_ln_bm <- brm(
  v_duration | mi(sd/sqrt(n_obs)) ~
    voice,
  data = p81_mono,
  prior = priors,
  family = lognormal(),
  seed = 9899,
  file = "./data/cache/p81_mono_ln_bm",
  control = list(adapt_delta = .9999, max_treedepth = 20),
  iter = iterations,
  save_mevars = TRUE
)
```

```{r luce1985}
l85_medial <- filter(stressed, study == "luce1985_medial")
l85_final <- filter(stressed, study == "luce1985_final")

priors <- c(
  prior(normal(0, 3.2), class = Intercept),
  prior(normal(0, 1), class = b, coef = voicevoiced),
  prior(normal(0, 0.1), class = sd),
  prior(normal(0, 0.1), class = sigma)
  # prior(lkj(2), class = cor)
)

l85_medial_ln_bm <- brm(
  v_duration | mi(sd/sqrt(n_obs)) ~
    voice + (1|pair),
  data = l85_medial,
  prior = priors,
  family = lognormal(),
  seed = 9899,
  file = "./data/cache/l85_medial_ln_bm",
  control = list(adapt_delta = .9999, max_treedepth = 20),
  iter = iterations,
  save_mevars = TRUE
)

l85_final_ln_bm <- brm(
  v_duration | mi(sd/sqrt(n_obs)) ~
    voice + (1|pair),
  data = l85_final,
  family = lognormal(),
  prior = priors,
  seed = 9899,
  file = "./data/cache/l85_final_ln_bm",
  control = list(adapt_delta = .9999, max_treedepth = 20),
  iter = iterations,
  save_mevars = TRUE
)
```

```{r davis1989}
d89 <- filter(stressed, study == "davis1989") %>%
  mutate(obs = 1:n())

priors <- c(
  prior(normal(0, 3), class = Intercept),
  prior(normal(0, 1), class = b, coef = voicevoiced),
  prior(normal(0, 0.1), class = sd),
  prior(normal(0, 0.1), class = sigma),
  prior(lkj(2), class = cor)
)

d89_ln_bm <- brm(
  v_duration | mi(sd/sqrt(n_obs)) ~
    voice + (1+voice|pair) + (1+voice|speaker),
  data = d89,
  prior = priors,
  family = lognormal(),
  seed = 9899,
  file = "./data/cache/d89_ln_bm",
  control = list(adapt_delta = .9999, max_treedepth = 20),
  iter = iterations
)
```

## Studies that provide only means (with no standard deviation)

Most studies only provide means, so Bayesian regressions are fitted to the means only.

```{r heffner1937}
h37 <- filter(stressed, study == "heffner1937")

priors <- c(
  prior(normal(0, 3), class = Intercept),
  prior(normal(0, 1), class = b, coef = voicevoiced),
  prior(normal(0, 0.1), class = sd),
  prior(normal(0, 0.1), class = sigma)
)

h37_ln_bm <- brm(
  v_duration ~
    voice + (1|pair),
  data = h37,
  prior = priors,
  family = lognormal(),
  seed = 9899,
  file = "./data/cache/h37_ln_bm",
  control = list(adapt_delta = .9999, max_treedepth = 15),
  iter = iterations
)
```

```{r housefairbanks1953}
hf53 <- filter(stressed, study == "housefairbanks1953")

priors <- c(
  prior(normal(0, 3), class = Intercept),
  prior(normal(0, 1), class = b, coef = voicevoiced),
  prior(normal(0, 0.1), class = sd),
  prior(normal(0, 0.1), class = sigma)
)

hf53_ln_bm <- brm(
  v_duration ~
    voice + (1|pair),
  data = hf53,
  prior = priors,
  family = lognormal(),
  seed = 9899,
  file = "./data/cache/hf53_ln_bm",
  control = list(adapt_delta = .9999, max_treedepth = 15),
  iter = iterations
)
```

```{r zimmerman1958}
z58 <- filter(stressed, study == "zimmerman1958")

priors <- c(
  prior(normal(0, 3), class = Intercept),
  prior(normal(0, 1), class = b, coef = voicevoiced),
  prior(normal(0, 0.1), class = sd),
  prior(normal(0, 0.1), class = sigma)
)

z58_ln_bm <- brm(
  v_duration ~
    voice + (1|pair),
  data = z58,
  prior = priors,
  family = lognormal(),
  seed = 9899,
  file = "./data/cache/z58_ln_bm",
  control = list(adapt_delta = .9999, max_treedepth = 15),
  iter = iterations
)
```

```{r petersonlehiste1960}
pl60 <- filter(stressed, study == "petersonlehiste1960")

priors <- c(
  prior(normal(0, 3), class = Intercept),
  prior(normal(0, 1), class = b, coef = voicevoiced),
  prior(normal(0, 0.1), class = sd),
  prior(normal(0, 0.1), class = sigma)
)

pl60_ln_bm <- brm(
  v_duration ~
    voice + (1|pair),
  data = pl60,
  prior = priors,
  family = lognormal(),
  seed = 9899,
  file = "./data/cache/pl60_ln_bm",
  control = list(adapt_delta = .9999, max_treedepth = 15),
  iter = iterations
)
```

```{r sharf1962}
s62_di <- filter(stressed, study == "sharf1962_di")
s62_mono <- filter(stressed, study == "sharf1962_mono")

priors <- c(
  prior(normal(0, 3), class = Intercept),
  prior(normal(0, 1), class = b, coef = voicevoiced),
  prior(normal(0, 0.1), class = sd),
  prior(normal(0, 0.1), class = sigma)
)

s62_di_ln_bm <- brm(
  v_duration ~
    voice + (1|pair),
  data = s62_di,
  prior = priors,
  family = lognormal(),
  seed = 9899,
  file = "./data/cache/s62_di_ln_bm",
  control = list(adapt_delta = .9999, max_treedepth = 15),
  iter = iterations
)

s62_mono_ln_bm <- brm(
  v_duration ~
    voice + (1|pair),
  data = s62_mono,
  prior = priors,
  family = lognormal(),
  seed = 9899,
  file = "./data/cache/s62_mono_ln_bm",
  control = list(adapt_delta = .9999, max_treedepth = 15),
  iter = iterations
)
```

```{r chen1970}
c70 <- filter(stressed, study == "chen1970")

priors <- c(
  prior(normal(0, 3), class = Intercept),
  prior(normal(0, 1), class = b, coef = voicevoiced),
  prior(normal(0, 0.1), class = sd),
  prior(normal(0, 0.1), class = sigma)
)

c70_ln_bm <- brm(
  v_duration ~
    voice + (1|pair),
  data = c70,
  prior = priors,
  family = lognormal(),
  seed = 9899,
  file = "./data/cache/c70_ln_bm",
  control = list(adapt_delta = .9999, max_treedepth = 15),
  iter = iterations
)
```

```{r mack1982}
m82 <- filter(stressed, study == "mack1982")

priors <- c(
  prior(normal(0, 3), class = Intercept),
  prior(normal(0, 1), class = b, coef = voicevoiced),
  prior(normal(0, 0.1), class = sd),
  prior(normal(0, 0.1), class = sigma)
)

m82_ln_bm <- brm(
  v_duration ~
    voice + (1|pair),
  data = m82,
  prior = priors,
  family = lognormal(),
  seed = 9899,
  file = "./data/cache/m82_ln_bm",
  control = list(adapt_delta = .9999, max_treedepth = 15),
  iter = iterations
)
```

```{r laeufer1992}
l92 <- filter(stressed, study == "laeufer1992")

priors <- c(
  prior(normal(0, 3), class = Intercept),
  prior(normal(0, 1), class = b, coef = voicevoiced),
  prior(normal(0, 0.1), class = sd),
  prior(normal(0, 0.1), class = sigma)
)

l92_ln_bm <- brm(
  v_duration ~
    voice + (1|pair),
  data = l92,
  prior = priors,
  family = lognormal(),
  seed = 9899,
  file = "./data/cache/l92_ln_bm",
  control = list(adapt_delta = .9999, max_treedepth = 15),
  iter = iterations
)
```

## Studies with raw data

```{r ko2018}
# k18 <- filter(ko18_data, phoneme %in% c("AE1", "EH1", "IH1")) %>%
#   mutate(
#     voice = factor(voice, levels = c("-voice", "+voice")),
#     phoneme = factor(phoneme, levels = c("AE1", "EH1", "IH1"))
#   )
#
# contrasts(k18$phoneme) <- "contr.sum"

priors <- c(
  prior(normal(0, 3), class = Intercept),
  prior(normal(0, 1), class = b, coef = voicePvoice),
  prior(normal(0, 1), class = b, coef = phoneme1),
  prior(normal(0, 1), class = b, coef = phoneme2),
  prior(normal(0, 1), class = b, coef = voicePvoice:phoneme1),
  prior(normal(0, 1), class = b, coef = voicePvoice:phoneme2),
  prior(cauchy(0, 0.1), class = sd),
  prior(cauchy(0, 0.1), class = sigma),
  prior(lkj(2), class = cor)
)

k18_ln_bm <- brm(
  duration ~
    voice *
    phoneme +
    (1+voice|subj),
  data = k18,
  prior = priors,
  family = lognormal(),
  seed = 9899,
  control = list(adapt_delta = 0.999, max_treedepth = 20),
  file = "./data/cache/k18_ln_bm"
)
```

```{r coretta2019}
contrasts(eng_durations$vowel) <- "contr.sum"
eng_durations_di <- filter(eng_durations, num_syl == "di")
eng_durations_mono <- filter(eng_durations, num_syl == "mono")

priors <- c(
  set_prior("normal(0, 3)", class = "Intercept"),
  set_prior("normal(0, 1)", class = "b", coef = "voicingvoiced"),
  set_prior("normal(0, 1)", class = "b", coef = "vowel1"),
  set_prior("normal(0, 1)", class = "b", coef = "vowel2"),
  set_prior("normal(0, 1)", class = "b", coef = "voicingvoiced:vowel1"),
  set_prior("normal(0, 1)", class = "b", coef = "voicingvoiced:vowel2"),
  set_prior("normal(0, 1)", class = "b", coef = "speech_rate_c"),
  set_prior("cauchy(0, 0.1)", class = "sd"),
  set_prior("lkj(2)", class = "cor"),
  set_prior("cauchy(0, 0.1)", class = "sigma")
)

c19_di_ln_bm <- brm(
  v1_duration ~
    voicing *
    vowel +
    speech_rate_c +
    (1 + voicing | speaker) +
    (1 | word),
  data = eng_durations_di,
  prior = priors,
  family = lognormal(),
  file = "./data/cache/c19_di_ln_bm",
  control = list(adapt_delta = .9999, max_treedepth = 15),
  iter = iterations
)

priors <- c(
  set_prior("normal(0, 3)", class = "Intercept"),
  set_prior("normal(0, 1)", class = "b", coef = "voicingvoiced"),
  set_prior("normal(0, 1)", class = "b", coef = "vowel1"),
  set_prior("normal(0, 1)", class = "b", coef = "vowel2"),
  set_prior("normal(0, 1)", class = "b", coef = "voicingvoiced:vowel1"),
  set_prior("normal(0, 1)", class = "b", coef = "voicingvoiced:vowel2"),
  set_prior("normal(0, 1)", class = "b", coef = "speech_rate_c"),
  set_prior("cauchy(0, 0.1)", class = "sd"),
  set_prior("lkj(2)", class = "cor"),
  set_prior("cauchy(0, 0.1)", class = "sigma")
)

c19_mono_ln_bm <- brm(
  v1_duration ~
    voicing *
    vowel +
    speech_rate_c +
    (1 + voicing | speaker) +
    (1 | word),
  family = lognormal(),
  data = eng_durations_mono,
  prior = priors,
  file = "./data/cache/c19_mono_ln_bm"
)
```

## Estimated effects with error

Now we make a data set with the estimated VE ratio and relative standard error from each study, to be used in the meta-analysis.

```{r estimated}
if (!file.exists("./data/cache/estimated_ln_tbl.rds")) {
  estimated_ln <- list()

  # These are in alphabetical order so that joining below does not complain
  # about different levels
  estimated_ln[["chen1970"]] <- fixef(c70_ln_bm)["voicevoiced",]
  estimated_ln[["coretta2019_di"]] <- fixef(c19_di_ln_bm)["voicingvoiced",]
  estimated_ln[["coretta2019_mono"]] <- fixef(c19_mono_ln_bm)["voicingvoiced",]
  estimated_ln[["davis1989"]] <- fixef(d89_ln_bm)["voicevoiced",]
  estimated_ln[["heffner1937"]] <- fixef(h37_ln_bm)["voicevoiced",]
  estimated_ln[["housefairbanks1953"]] <- fixef(hf53_ln_bm)["voicevoiced",]
  estimated_ln[["ko2018"]] <- fixef(k18_ln_bm)["voicePvoice",]
  estimated_ln[["laeufer1992"]] <- fixef(l92_ln_bm)["voicevoiced",]
  estimated_ln[["luce1985_final"]] <- fixef(l85_final_ln_bm)["voicevoiced",]
  estimated_ln[["luce1985_medial"]] <- fixef(l85_medial_ln_bm)["voicevoiced",]
  estimated_ln[["mack1982"]] <- fixef(m82_ln_bm)["voicevoiced",]
  estimated_ln[["petersonlehiste1960"]] <- fixef(pl60_ln_bm)["voicevoiced",]
  estimated_ln[["port1981_di"]] <- fixef(p81_di_ln_bm)["voicevoiced",]
  estimated_ln[["port1981_mono"]] <- fixef(p81_mono_ln_bm)["voicevoiced",]
  estimated_ln[["sharf1962_di"]] <- fixef(s62_di_ln_bm)["voicevoiced",]
  estimated_ln[["sharf1962_mono"]] <- fixef(s62_mono_ln_bm)["voicevoiced",]
  estimated_ln[["zimmerman1958"]] <- fixef(z58_ln_bm)["voicevoiced",]

  estimated_ln_tbl <- plyr::ldply(estimated_ln, .id = "study") %>%
    mutate(
      source = factor("original", levels = c("original", "meta-analysis")),
      study = factor(study)
    ) %>%
    left_join(y = studies)

  saveRDS(estimated_ln_tbl, file = "./data/cache/estimated_ln_tbl.rds")
}
```

# Meta-analysis

We can now fit a meta-analytical model using a Bayesian linear regression with measurement error (added with `se(Est.Error)` in the left-hand side of the model formula).

```{r meta-bm}
priors <- c(
  prior(normal(0, 3), class = Intercept),
  prior(normal(0, 1), class = b, coef = syl_posnonMfinal),
  prior(normal(0, 0.1), class = sd)
)

meta_ln_bm <- brm(
  Estimate | se(`Est.Error`) ~ syl_pos + (1 | study),
  data = estimated_ln_tbl,
  prior = priors,
  control = list(adapt_delta = 0.9999, max_treedepth = 15),
  seed = 9899,
  file = "./data/cache/meta_ln_bm"
)

summary(meta_ln_bm)

meta_ln_bm_est <- fixef(meta_ln_bm)["Intercept","Estimate"]
meta_ln_bm_q2.5 <- fixef(meta_ln_bm)["Intercept","Q2.5"]
meta_ln_bm_q97.5 <- fixef(meta_ln_bm)["Intercept","Q97.5"]

post_ln_nonfi <- (c(posterior_samples(meta_ln_bm, pars = "b_Intercept")) +
  posterior_samples(meta_ln_bm, pars = "b_syl_posnonMfinal"))

meta_ln_bm_syl_q2.5 <- quantile(post_ln_nonfi$b_syl_posnonMfinal, c(0.025))
meta_ln_bm_syl_q97.5 <- quantile(post_ln_nonfi$b_syl_posnonMfinal, c(0.975))
meta_ln_bm_syl_q50 <- (meta_ln_bm_syl_q2.5 + meta_ln_bm_syl_q97.5) / 2
```

## Marginal posterior distributions

```{r marginal-post}
marginal_ln_post <- posterior_samples(meta_ln_bm, pars = "b_") %>%
  mutate(
    final = exp(b_Intercept),
    `non-final` = exp(b_Intercept + b_syl_posnonMfinal)
  ) %>%
  select(final:`non-final`) %>%
  pivot_longer(everything(), names_to = "position", values_to = "posterior")

marginal_ln_post %>%
  ggplot(aes(posterior, fill = position)) +
  geom_vline(aes(xintercept = 1)) +
  geom_density(colour = NA, alpha = 0.8) +
  scale_fill_brewer(
    type = "qual", palette = "Paired"
  ) +
  labs(
    title = "Meta-analytical marginal posterior distributions of the voicing effect",
    x = "Difference in vowel duration (ratio)",
    fill = "Syllable position",
    y = element_blank()
  ) +
  theme(legend.position = "top")
```

## Shrinkage

```{r origin-1, include=FALSE}
estimated_ln_tbl %>%
  ggplot(aes(reorder(study_ref, Estimate), exp(Estimate), linetype = syl_pos)) +
  geom_hline(aes(yintercept = 1), colour = "grey") +
  scale_x_discrete() +
  geom_point(size = 2, position = position_dodge(width = 0.8)) +
  geom_errorbar(aes(ymin = exp(Q2.5), ymax = exp(Q97.5)), width = 0, position = position_dodge(width = 0.8)) +
  scale_color_brewer(type = "qual", palette = "Dark2", breaks = c("original", "meta-analysis")) +
  scale_y_continuous(breaks = seq(0.5, 2.5, by = 0.25)) +
  scale_linetype_discrete(labels = c("final", "non-final")) +
  labs(
    # caption = "The shaded areas indicate the 95% CI of the meta-analytical posterior\nof the voicing effect in final (light blue) and penultimate (dark blue) position.",
    title = "Studies' estimates of the Voicing Effect",
    y = "Difference in vowel duration (ratio)",
    x = "Study",
    linetype = "Syllable position",
    colour = "Source"
  ) +
  coord_flip() +
  theme(panel.grid.major.y = element_blank(), panel.grid.minor.y = element_blank())


ggsave("./img/origin-1.png", width = 7, height = 4.5)
```

```{r origin-2, include=FALSE}
estimated_ln_tbl %>%
  ggplot(aes(reorder(study_ref, Estimate), exp(Estimate), linetype = syl_pos)) +
  geom_hline(aes(yintercept = 1), colour = "grey") +
  scale_x_discrete() +
  annotate("rect", xmin = 4.5, xmax = Inf, ymin = exp(meta_ln_bm_q2.5), ymax = exp(meta_ln_bm_q97.5), alpha = 0.5, fill = "#a6cee3") +
  geom_point(size = 2, position = position_dodge(width = 0.8)) +
  geom_errorbar(aes(ymin = exp(Q2.5), ymax = exp(Q97.5)), width = 0, position = position_dodge(width = 0.8)) +
  scale_color_brewer(type = "qual", palette = "Dark2", breaks = c("original", "meta-analysis")) +
  scale_y_continuous(breaks = seq(0.5, 2.5, by = 0.25)) +
  scale_linetype_discrete(labels = c("final", "non-final")) +
  labs(
    # caption = "The shaded areas indicate the 95% CI of the meta-analytical posterior\nof the voicing effect in final (light blue) and penultimate (dark blue) position.",
    title = "Studies' estimates of the Voicing Effect",
    y = "Difference in vowel duration (ratio)",
    x = "Study",
    linetype = "Syllable position",
    colour = "Source"
  ) +
  coord_flip() +
  theme(panel.grid.major.y = element_blank(), panel.grid.minor.y = element_blank())


ggsave("./img/origin-2.png", width = 7, height = 4.5)
```

```{r origin-3}
estimated_ln_tbl %>%
  ggplot(aes(reorder(study_ref, Estimate), exp(Estimate), linetype = syl_pos)) +
  geom_hline(aes(yintercept = 1), colour = "grey") +
  scale_x_discrete() +
  annotate("rect", xmin = 4.5, xmax = Inf, ymin = exp(meta_ln_bm_q2.5), ymax = exp(meta_ln_bm_q97.5), alpha = 0.5, fill = "#a6cee3") +
  annotate("rect", xmin = -Inf, xmax = 4.5, ymin = exp(meta_ln_bm_syl_q2.5), ymax = exp(meta_ln_bm_syl_q97.5), alpha = 0.5, fill = "#1f78b4") +
  geom_point(size = 2, position = position_dodge(width = 0.8)) +
  geom_errorbar(aes(ymin = exp(Q2.5), ymax = exp(Q97.5)), width = 0, position = position_dodge(width = 0.8)) +
  scale_color_brewer(type = "qual", palette = "Dark2", breaks = c("original", "meta-analysis")) +
  scale_y_continuous(breaks = seq(0.5, 2.5, by = 0.25)) +
  scale_linetype_discrete(labels = c("final", "non-final")) +
  labs(
    # caption = "The shaded areas indicate the 95% CI of the meta-analytical posterior\nof the voicing effect in final (light blue) and penultimate (dark blue) position.",
    title = "Studies' estimates of the Voicing Effect",
    y = "Difference in vowel duration (ratio)",
    x = "Study",
    linetype = "Syllable position",
    colour = "Source"
  ) +
  coord_flip() +
  theme(panel.grid.major.y = element_blank(), panel.grid.minor.y = element_blank())

ggsave("./img/origin-3.png", width = 7, height = 4.5)
```

```{r studies-shrunk}
# Code adapted from Nicenboim et al. 2018

# We need this since the model includes `num_syl`
post_nonf <- posterior_samples(meta_ln_bm, pars = "b_syl_posnonMfinal")

studies_shrunk <- (c(posterior_samples(meta_ln_bm, pars = "b_Intercept")) +
  posterior_samples(meta_ln_bm, pars = "r_study")) %>%
  bind_cols(post_nonf) %>%
  # The following mutate sums the random coefficients of the studies with
  # `num_syl` == "non-final" to the posterior samples of the non_syl term
  mutate(
    `r_study[davis1989,Intercept]` = `r_study[davis1989,Intercept]` + b_syl_posnonMfinal,
    `r_study[sharf1962_di,Intercept]` = `r_study[sharf1962_di,Intercept]` + b_syl_posnonMfinal,
    `r_study[coretta2019_di,Intercept]` = `r_study[coretta2019_di,Intercept]` + b_syl_posnonMfinal,
    `r_study[port1981_di,Intercept]` = `r_study[port1981_di,Intercept]` + b_syl_posnonMfinal,
  ) %>%
  select(-b_syl_posnonMfinal) %>%
  summarise_all(list(~list(c(
    mean(.),
    quantile(., probs = c(.025, 0.975)),
    sd(.)
  )))) %>%
  unnest(cols = c(`r_study[chen1970,Intercept]`, `r_study[coretta2019_di,Intercept]`,
    `r_study[coretta2019_mono,Intercept]`, `r_study[davis1989,Intercept]`,
    `r_study[heffner1937,Intercept]`, `r_study[housefairbanks1953,Intercept]`,
    `r_study[ko2018,Intercept]`, `r_study[laeufer1992,Intercept]`,
    `r_study[luce1985_final,Intercept]`, `r_study[luce1985_medial,Intercept]`,
    `r_study[mack1982,Intercept]`, `r_study[petersonlehiste1960,Intercept]`,
    `r_study[port1981_di,Intercept]`, `r_study[port1981_mono,Intercept]`,
    `r_study[sharf1962_di,Intercept]`, `r_study[sharf1962_mono,Intercept]`,
    `r_study[zimmerman1958,Intercept]`)) %>%
  transpose() %>%
  setNames(c("Estimate", "Q2.5", "Q97.5", "Est.Error")) %>%
  map_df(unlist) %>%
  mutate(
    study = estimated_ln_tbl$study,
    source = factor("meta-analysis", levels = c("original", "meta-analysis"))
  ) %>%
  left_join(y = studies)
```

```{r origin-shrunk-plot, include=TRUE, fig.cap = "Estimated voicing effect from the original source and from the meta-analysis.", out.extra = "width=\\textwidth"}
bind_rows(estimated_ln_tbl, studies_shrunk) %>%
  mutate(source = factor(source, levels = c("meta-analysis", "original"))) %>%
  ggplot(aes(reorder(study_ref, Estimate), Estimate, colour = source, linetype = syl_pos)) +
  geom_hline(aes(yintercept = 0), colour = "grey") +
  scale_x_discrete() +
  annotate("rect", xmin = 4.5, xmax = Inf, ymin = meta_ln_bm_q2.5, ymax = meta_ln_bm_q97.5, alpha = 0.5, fill = "#a6cee3") +
  annotate("rect", xmin = -Inf, xmax = 4.5, ymin = meta_ln_bm_syl_q2.5, ymax = meta_ln_bm_syl_q97.5, alpha = 0.5, fill = "#1f78b4") +
  geom_point(size = 2, position = position_dodge(width = 0.8)) +
  geom_errorbar(aes(ymin = Q2.5, ymax = Q97.5), width = 0, position = position_dodge(width = 0.8)) +
  scale_color_brewer(type = "qual", palette = "Dark2", breaks = c("original", "meta-analysis")) +
  scale_y_continuous(breaks = c(-0.4, -0.2, 0, 0.2, 0.4, 0.6, 0.8, 1)) +
  scale_linetype_discrete(labels = c("final", "non-final")) +
  labs(
    caption = "The shaded areas indicate the 95% CI of the meta-analytical posterior\nof the voicing effect in final (light blue) and penultimate (dark blue) position.",
    y = "Difference in vowel duration (log-odds)",
    x = "Study",
    linetype = "Syllable position",
    colour = "Source"
  ) +
  coord_flip() +
  theme(panel.grid.major.y = element_blank(), panel.grid.minor.y = element_blank())
```

# Bias

Potential biases in the meta-analytical estimate of the VE ratio are assessed with the funnel plot method [@light1984].
For a review of the method see @sterne2005 and @egger1997, for a critique @lau2006.

In presence of bias, the dots in the funnel plot are not symmetrically distributed around the meta-analytical mean (the gray vertical line).

```{r funnel-sd, include=TRUE}
e.max <- by(estimated_ln_tbl$Est.Error, estimated_ln_tbl$syl_pos, range)[[1]][2]
se_range <- seq(0, e.max, by = 0.001)
ci <- tibble(
  x_seq = se_range,
  ci_lo = meta_ln_bm_est - 1.96 * se_range,
  ci_up = meta_ln_bm_est + 1.96 * se_range
)

estimated_ln_tbl %>%
  filter(syl_pos == "final") %>%
  ggplot(aes(Estimate, Est.Error)) +
  geom_line(aes(y = x_seq, x = ci_up), data = ci) +
  geom_line(aes(y = x_seq, x = ci_lo), data = ci) +
  geom_vline(aes(xintercept = 0), colour = "grey") +
  geom_vline(aes(xintercept = meta_ln_bm_est)) +
  annotate("rect", ymin = -Inf, ymax = Inf, xmin = meta_ln_bm_q2.5, xmax = meta_ln_bm_q97.5, alpha = 0.5, fill = "#a6cee3") +
  geom_point(size = 3) +
  labs(
    title = "By-study funnel plot of the estimates against their standard deviation",
    x = "Difference in vowel duration (log-odds)"
  ) +
  scale_y_reverse()

ggsave("./img/funnel-sd.png", width = 7, height = 4.5)
```

There is some indication of bias towards larger values (two studies with x higher than 0.6).

```{r funnel-precision, include=TRUE, fig.cap = "By-study funnel plot showing the estimate against the precision. The vertical thick and dashed lines are the meta-analytical means of the effect in final and penultimate position."}
e.min <- by(estimated_ln_tbl$Est.Error, estimated_ln_tbl$syl_pos, range)[[1]][1]
e.max <- by(estimated_ln_tbl$Est.Error, estimated_ln_tbl$syl_pos, range)[[1]][2]
se_range <- seq(e.min, e.max, by = 0.001)
ci <- tibble(
  y_seq = 1/(seq(e.min, e.max, by = 0.001)^2),
  ci_lo = meta_ln_bm_est - 1.96 * se_range,
  ci_up = meta_ln_bm_est + 1.96 * se_range
)

# plot(seq(0, 30, by = 1) * 1.96, 1/seq(0, 30, by = 1))

estimated_ln_tbl %>%
  mutate(precision = 1/(Est.Error^2)) %>%
  filter(syl_pos == "final") %>%
  ggplot(aes(Estimate, precision)) +
  geom_vline(aes(xintercept = 0), colour = "grey") +
  geom_vline(aes(xintercept = meta_ln_bm_est)) +
  annotate("rect", ymin = -Inf, ymax = Inf, xmin = meta_ln_bm_q2.5, xmax = meta_ln_bm_q97.5, alpha = 0.5, fill = "#a6cee3") +
  geom_point(size = 3) +
  scale_x_continuous(breaks = seq(-1, 1, 0.2)) +
  labs(
    x = "Difference in vowel duration (log-odds)",
    shape = "Syllable position"
  ) +
  geom_line(aes(y = y_seq, x = ci_up), data = ci) +
  geom_line(aes(y = y_seq, x = ci_lo), data = ci)
```

# References
